---
title: "Final.project"
author: "Samira Bechi"
date: "2023-11-24"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## INTRODUCTION
[link for the Airbnb data](https://www.kaggle.com/datasets/PromptCloudHQ/airbnb-property-data-from-texas)
Airbnb is an online platform that facilitates the booking and rental of lodging, primarily homes or apartments, on a short-term basis. Founded in 2008, Airbnb has grown into a global marketplace that connects people looking to rent their homes with travelers seeking unique and often more affordable accommodations than traditional hotels. The platform allows hosts to list their properties, providing details about the space, amenities, and pricing, while guests can browse listings and book accommodations for a specified period. 
I've heard many good things about this business but have not seen a market study. Today, I decided to work with this data (Airbnb property data from Texas from 2008 to 2017). This dataset includes the average night price, number of bedrooms, location, the year listed, and description, which aligns with my interest in studying the market. At the end of our analyses, we will be able to respond to multiple questions such as how many listings is there in Texas, how Airbnb has grown in Texas from 2008 to 2017, what city is the most popular with Airbnb; what is the average per night, per year, per bedroom; what number of the bedroom is the most popular?
Through our analysis, we will find approaches to our findings. Our data will give a visualization and give people an idea of this business.

## LIBRARY USED

```{r library}
library(kableExtra) # used to construct Complex Table for data
library(tidyverse) #using this to tidy up the data
library(dplyr) #using this for data manipulation
library(ggplot2) #using for visual data
```

## UNFILTERED DATA

```{r load_data}
load("Data.RData")
glimpse(Airbnb_Texas)
```

## FORMATING DATA
```{r format_data}
#removing all unecessary data
Airbnb_Texas$description <- NULL
Airbnb_Texas$latitude <- NULL
Airbnb_Texas$longitude <- NULL
Airbnb_Texas$url <- NULL
Airbnb_Texas$title <- NULL
# separate year and moth to be able to access data easily
date <- strsplit(x = as.character(Airbnb_Texas$date_of_listing) ,split = " ",fixed = T)
Airbnb_Texas$month <- as.factor(sapply(X = date,"[[",1))
Airbnb_Texas$year <- as.factor(sapply(X = date,"[[",2))
Airbnb_Texas$date_of_listing <- NULL
# remove $ sign and convert the average per night to a numeric 
Airbnb_Texas$average_rate_per_night <-substr(x = Airbnb_Texas$average_rate_per_night,start = 2,stop = 10)
Airbnb_Texas$average_rate_per_night <- as.numeric(Airbnb_Texas$average_rate_per_night)
# consider a studio as a 1 bedroom and change it to numeric 
Airbnb_Texas$bedrooms_count[Airbnb_Texas$bedrooms_count=="Studio"] <- "1"
Airbnb_Texas$bedrooms_count <- as.numeric(Airbnb_Texas$bedrooms_count)
glimpse(Airbnb_Texas)
```
## DATA KEY TERMS
**average_rate_per_night** cost of the room per night
**bedrooms_count** number of bedroom
**city** location of the airbnb
**date_of_listing** date when airbnb was available on platform
## 1- Total Airbnb from 2008 to 2017
From 2008 to 2017 exactly 18,259 were created in Texas.
## 2- Number of houses by year
We are taking our formatted data to display the number of houses per year. This will help us determine what year had the most additional Airbnbs and if there was an increase or decrease in the number of Airbnbs added yearly.
```{r}
ggplot(Airbnb_Texas %>% group_by(year) %>% count(), aes(year,n,fill=year,color=year)) +
   geom_col() +
   labs(title=" Number of houses by year in Texas",subtitle="By Year")
```

## 4- Top 10 city with most airbnb 
```{r}
Airbnb_Texas %>% group_by(city) %>% count() %>% arrange(desc(n)) %>% kbl() %>% row_spec(1:10, background = "pink") %>% kable_paper()  %>% scroll_box(height = "300px")
```

## 5- Average rate per night per month 
```{r}
# Create an ordered factor for 'month'
order_of_months <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
Airbnb_Texas$month <- factor(Airbnb_Texas$month, levels = order_of_months)

# Plot
ggplot(Airbnb_Texas %>% 
         group_by(month) %>%
         summarize(mean_rate = mean(average_rate_per_night, na.rm = TRUE)),
       aes(month, mean_rate, fill = month, color = month)) +
  geom_col() +
  labs(title = "Average Rate Per Month") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
## 6- Average rate per night per year
```{r}
ggplot(Airbnb_Texas %>% 
         group_by(year) %>%
         summarize(mean_rate = mean(average_rate_per_night, na.rm = TRUE)),
       aes(x = year, y = mean_rate, group = 1)) +
  geom_line() +
  labs(title = "Average Rate Over Years")

```
## 7- Number of bedrooms 
```{r}
Airbnb_Texas$bedrooms_count[Airbnb_Texas$bedrooms_count=="Studio"] <- "1"
Airbnb_Texas$bedrooms_count <- as.numeric(Airbnb_Texas$bedrooms_count)
 ggplot(Airbnb_Texas , aes(x = bedrooms_count, fill = as.factor(bedrooms_count))) +
  geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.7) +
  scale_fill_manual(values = rainbow(length(unique(Airbnb_Texas$bedrooms_count)))) +
   labs(title = "Number of Bedrooms")

```
## 8-Average rate night per bedroom
```{r}
Airbnb_Texas %>% 
  group_by(bedrooms_count) %>%
  summarize(mean_rate = mean(average_rate_per_night, na.rm = TRUE))
```
## Conclusion
From 2008 to 2017, 18259 were listed in Texas, but massive growth started around 2011. When we look at the top 10 cities with the most listings, we realize they are the big cities of Texas with many visitors, international airports, attractions... The average rate per night has considerably grown over the years. One bedroom is the most popular listing because it's the cheapest average night. Overall, Airbnb is a trendy business in Texas, and the questions we answered should help investors direct themselves into this business.
